---
title: "paired_t_test"
output: github_document
---


```{r}
library(tidyverse)
library(dplyr)
```


```{r}
#find sample size, mean, and standard deviation for each group
data %>%
  group_by(group) %>%
  summarise(
    count = n(),
    mean = mean(score),
    sd = sd(score)
  )
```

```{r}
# Create boxplots using the boxplot() function in R to view the distribution of scores for the pre and post groups:
boxplot(score~group,
  data=data,
  main="Test Scores by Group",
  xlab="Group",
  ylab="Score",
  col="steelblue",
  border="black"
)
```





**Conduct a Paired t-test**

Before we conduct the paired t-test, we should check that the distribution of differences is normally (or approximately normally) distributed. To do so, we can create a new vector defined as the difference between the pre and post scores, and perform a shapiro-wilk test for normality on this vector of values:
```{r}
#define new vector for difference between post and pre scores
differences <- with(data, score[group == "post"] - score[group == "pre"])

#perform shapiro-wilk test for normality on this vector of values
shapiro.test(differences)
```
If p-value is greater than alpha = 0.05. -> We fail to reject the null hypothesis that our data is normally distributed. 
This means we can now proceed to conduct the paired t-test.



```{r}
t.test(x, y, data = data, paired = TRUE, alternative = 'two.sided')
```


